---
title: "AETree"
author: "datadavidz"
date: "August 29, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(caTools)
library(rpart)
library(rpart.plot)
library(caret)
library(ROCR)
```

## Decision Tree

Supreme court cases from 1994 to 2001.  Same justices in this time period: Breyer, Ginsburg, Kennedy, O’Connor, Rehnquist (Chief Justice), Scalia, Souter, Stevens, Thomas.  Focus on predicting Judge Stevens decisions was a moderate but became 
more liberal over the time period.

```{r}
stevens <- read.csv("Data/stevens.csv")
```

###Dependent variable 
**Reverse:**  Whether Judge Stevens decided to reverse the lower court decision (1 = Yes, 0 = No)

###Independent variables
Circuit: Circuit court of origin (1st-11th, DC, FED)
Issue: Issue area of case (EconomicActivity, JudicialPower etc.)
Petitioner: Type of petitioner (Business, City, Employee etc.)
Respondent:  Type of respondent (Business, City etc.)
LowerCourt: Ideological direction of lower court decision (Liberal or Conservative)
Unconst: Whether petitioner argued the law/practice was unconstitutional (1 = Yes, 0 = No)

```{r stevens_CART}
set.seed(3000)

#Split into training and testing datasets
spl <- sample.split(stevens$Reverse, SplitRatio=0.7)
Train <- subset(stevens, spl==TRUE)
Test <- subset(stevens, spl==FALSE)

#minbucket is lower bound for number of points in each subset
StevensTree <- rpart(Reverse ~ Circuit + Issue + Petitioner + Respondent + LowerCourt + Unconst, data=Train, method="class", minbucket=25)
```

## Stevens Tree Predicting Decisions

```{r stevens_plot, echo=FALSE}
#Uses the rpart.plot library
prp(StevensTree)
```

Predictive performance

```{r}
PredictCART <- predict(StevensTree, newdata=Test, type="class")
#confusion matrix and statistics from caret package
confusionMatrix(Test$Reverse, PredictCART)
```

ROCR Assessment of Predictive Performance

```{r}
#Varying the threshold for a positive prediction to see effect on True Positive Rate vs. False Positive Rate
PredictROC <- predict(StevensTree, newdata=Test)
pred <- prediction(PredictROC[,2], Test$Reverse)
perf <- performance(pred, "tpr", "fpr")
plot(perf)

auc <- as.numeric(performance(pred, "auc")@y.values)
```

## Cross Validation to select rpart tuning parameter

```{r}
#For classification, caret requires outcome to be factor
Train$Reverse <- as.factor(Train$Reverse)
Test$Reverse <- as.factor(Test$Reverse)

#Hold out ~40 samples per fold
numFolds <- trainControl(method="cv", number=10)

#For tuning CART it is better to use the complexity parameter (cp)
#Smaller cp gives a bigger tree (possibly overfitting)
cpGrid <- expand.grid(.cp=seq(0.01, 0.5, 0.01))

train(Reverse ~ Circuit + Issue + Petitioner + Respondent + LowerCourt + Unconst, data=Train, method="rpart", 
      trControl=numFolds, tuneGrid=cpGrid)
```

Build tree using cp value determined by cross-validation

```{r}
StevensTreeCV <- rpart(Reverse ~ Circuit + Issue + Petitioner + Respondent + LowerCourt + Unconst, data=Train, method="class", cp=0.18)
PredictCV <- predict(StevensTreeCV, newdata=Test, type="class")
confusionMatrix(Test$Reverse, PredictCV)
```

